--- 
title: "Penguins size exploration Troubleshooting Activity, Michelle Geldin"
output: html_document
---

```{r setup, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(palmerpenguins)
library(tidyverse) 
library(kableExtra) # Note: you probably need to install this package
```

### R Activity 1: Troubleshoot this code

In this activity, troubleshoot the code. Each line with a `# FIX ME` at the end has a single error that needs to be fixed. We recommend trying to run code line-by-line as you fix each to ensure that it's working, then move on to the next line. 

## 1. Penguins data exploratory visualization

In this section, troubleshoot the code to prepare 2 customized ggplot graphs using the `penguins` dataset from the `palmerpenguins` R package: 

- A scatterplot of flipper length vs. body mass, faceted (split up) by island
- A histogram of bill lengths for all gentoo penguins in the data

#### Scatterplot (flipper length & body mass)

```{r}
ggplot(data = penguins, 
       aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(aes(color = "species", size = 2)) + 
  scale_color_manual(values = c("gray20","orange","cyan4"), 
                     name = "Penguin species:") +
  scale_x_continuous(breaks=c(170, 200, 230)) +
  theme_bw() +
  facet_wrap(~ island) +
  labs(y = "Body mass (g)",
       x = "Flipper length (mm)", 
       title = "Palmer Archipelago adult nesting penguin size by island",
       subtitle = "Data from Kristen Gorman, Palmer Station LTER (2007 - 2009)") +
  theme(legend.position = "bottom")
```
#### Histogram of Gentoo bill lengths

```{r}
gentoo <- penguins %>%
  filter(species == "Gentoo")

ggplot(data = gentoo, aes(x = bill_length_mm)) +
  geom_histogram(fill = "goldenrod3", 
                 color = "goldenrod4",
                 bins = 20) +
  theme_minimal() +
  labs(x = "Bill length (mm)",
       y = "Frequency",
       title = "Gentoo penguin bill length distribution",
       subtitle = "Adults recorded in Palmer Archipelago, 2007 - 2009")
```
## 2. Penguins summary statistics

In the following examples, troubleshoot the code to prepare several tables of summary statistics. You need to have the `kableExtra` R package installed (run `install.packages("kableExtra")` in the Console) and attached for the tables to be created successfully. 

```{r}
f_penguin_mass <- penguins %>% 
  filter(sex == "female") %>%
  group_by(species) %>%
  summarize(mean_mass_g = mean(body_mass_g, na.rm = TRUE),
            max_mass_g = max(body_mass_g),
            sample_size = n())

f_penguin_mass %>% 
  kable(col.names = c("Species",
                      "Mean body mass (g)", 
                      "Max body mass (g)",
                      "Sample size (n)")) %>% 
  kable_styling(full_width = FALSE)
```

## 3. Penguin wrangling

The goal of this code is to: 

- Starting from `penguins`...
- Only keep observations for Adelie and Chinstrap penguins from 2008
- Select only columns for species, body mass, and flipper length
- Add a new column with body mass converted to kilograms
- Remove the body mass (grams) column
- Rename the flipper length column to 'flip_mm'
- Keep the top 5 largest (by body mass, in kg) penguins for both species
- Create a table of the outcome

```{r}
adelie_chinstrap_mass_top5 <- penguins %>%
  filter(species %in% c("Adelie", "Chinstrap"), year == 2008) %>%
  select(species, body_mass_g, flipper_length_mm) %>%
  mutate(mass_kg = body_mass_g / 1000) %>%
  select (-body_mass_g) %>%
  rename(flip_mm = flipper_length_mm) %>% 
  group_by(species) %>% 
  slice_max(order_by = mass_kg, n = 5)
  
adelie_chinstrap_mass_top5 %>% 
  kable(col.names = c("Species",
                      "flip_mm ",
                      "Body mass (kg)")) %>%  
  kable_styling(full_width = FALSE)
```

## END PENGUIN TROUBLE
